---
title: "DiCoLo_Figure_Code"
author: "Ruiqi Li"
date: '2026-01-13'
output: pdf_document
---
# Load library
```{r}
library("Seurat")
library("ggplot2")
library("patchwork")
devtools::install_github("KlugerLab/DiCoLo")
library("DiCoLo")

# source("../DiCoLo/vignettes/DiCoLo_functions.R")
source("./simulation_functions.R")
source("./plotting_functions.R")
```

# Figure 2A
Load pbmc10k data
```{r}
dir.path = "DiCoLo_data"
data.path = file.path(dir.path,"pbmc10k")

data_S_merge = readRDS(file.path(data.path,"pbmc10k.rds"))
DimPlot(data_S_merge, group.by = 'batch', shuffle = TRUE) + NoAxes() + 
  ggtitle("CD14+ Monocytes")
```
Select cell neighborhoods and simulate co-localized gene modules
```{r}
res.path = file.path(data.path,"simulation_parameters")
if(!dir.exists(res.path)){
  dir.create(res.path,recursive = TRUE)
}
```
```{r}
params_ls = list()
for(i in 1:2){
  file_path = file.path(res.path,sprintf("params%d.rds",i))
  if(!file.exists(file_path)){
diff.pct = 0.7; zinb_prob = 0.4
seed_ls = c(42,1208)
n_neighbors_ls = c(4,4)
sample_ls_short = paste0(sample_ls,"_injected","_",n_neighbors_ls)
ngene_per_neighbors_ls = c(25,15,15,20)
params_ls = lapply(1:2, function(i){
  n_neighbors = n_neighbors_ls[i]
  if(n_neighbors==0){
    return(NULL)
  }else{
    set.seed(seed_ls[i])
    ncell_per_neighbors_range = seq(0.03,0.1,0.01)
    ncell_per_neighbors_range = floor(ncol(data_S_ls[[i]]) * ncell_per_neighbors_range)
    ncell_per_neighbors_ls = sample(ncell_per_neighbors_range,n_neighbors)
    gene_params = Generate_gene_parameters(data_S_ls[i][[1]], data_S_ls[-i][[1]], 
                                           assay = "RNA", layer = "counts",
                                           n_neighbors = n_neighbors,
                                           ngene_per_neighbors_ls = ngene_per_neighbors_ls, 
                                           seed = seed_ls[i],
                                           mean_quantile = 0.9)
    return(list(ncell_per_neighbors_ls = ncell_per_neighbors_ls,
                seed = seed_ls[i],
                zinb_prob = zinb_prob,
                diff.pct = diff.pct,
                gene_params = gene_params))
  }
})

# Define common group
set.seed(seed_ls[1])
group_n = 2
group_id = sample(1:n_neighbors_ls[1],group_n)
params_ls[[1]]$ncell_per_neighbors_ls[group_id] = params_ls[[1]]$ncell_per_neighbors_ls[group_id] * 2
params_ls[[2]]$ncell_per_neighbors_ls[group_id] = params_ls[[1]]$ncell_per_neighbors_ls[group_id]
params_ls[[2]]$gene_params$selected_gene_groups[group_id] = params_ls[[1]]$gene_params$selected_gene_groups[group_id]

# re-name gene group
for(i in 1:2){
  a = names(params_ls[[i]]$gene_params$selected_gene_groups)
  a = ifelse(a %in% group_id, paste0("c_",a), paste0(i,"_",a))
  names(params_ls[[i]]$gene_params$selected_gene_groups) = a
  names(params_ls[[i]]$ncell_per_neighbors_ls) = a
}

# save params
lapply(1:2,function(i){
  saveRDS(params_ls[[i]],file = file.path(res.path,sprintf("params%d.rds",i)))
})
  }
  else{
    params_ls[[i]] = readRDS(file = file_path)
  }
}

injected_genes_ls = lapply(params_ls,function(params){
  if(is.null(params)){return(NULL)}
  with(as.list(params),{
    return(setNames(unlist(mapply(function(genes, val) rep(val, length(genes)), 
                                  gene_params$selected_gene_groups, names(gene_params$selected_gene_groups), SIMPLIFY = FALSE)),
                    unlist(gene_params$selected_gene_groups)))
  })
})
injected_genes = unlist(injected_genes_ls)
injected_genes = injected_genes[!duplicated(names(injected_genes))]
# define module_cols
color_m = setNames(brewer.pal(9,"Set1")[1:length(unique(injected_genes))],unique(injected_genes))
color_m = gsub("#FFFF33","#F781BF",color_m)

gene_params_merged = Map(c,params_ls[[1]]$gene_params,params_ls[[2]]$gene_params)
tmp_genegroup = gene_params_merged$selected_gene_groups
tmp_genegroup = tapply(seq_along(tmp_genegroup), names(tmp_genegroup), function(i) unique(unlist(tmp_genegroup[i])), simplify = FALSE)
gene_params_merged$selected_gene_groups = tmp_genegroup
gene_params_merged[2:5] = lapply(gene_params_merged[2:5],function(x) mean(x))
other_params_merged = Map(c,params_ls[[1]][1:4],params_ls[[2]][1:4])
other_params_merged$ncell_per_neighbors_ls = other_params_merged$ncell_per_neighbors_ls[!duplicated(names(other_params_merged$ncell_per_neighbors_ls))]
other_params_merged[2:4] = lapply(other_params_merged[2:4],function(x) unique(x))
params_merged = c(other_params_merged,list(gene_params = gene_params_merged))
params_merged$ncell_per_neighbors_ls = params_merged$ncell_per_neighbors_ls[names(params_merged$gene_params$selected_gene_groups)]
```
```{r}
# modify gene expression
condition_vec = as.factor(data_S_merge$batch)
data_S_merge_test = with(as.list(params_merged), {
  srat_injected <- Inject_local_signals_v2(data_S_merge, condition_vec = condition_vec, 
                                           assay = "RNA",
                                           layer = "counts",
                                           reduc = "pca", npc = 10, 
                                           min_knn = max(ncell_per_neighbors_ls),
                                           # min_knn = sum(sort(ncell_per_neighbors_ls,decreasing = T)[1:2]), 
                                           ncell_per_neighbors_ls = ncell_per_neighbors_ls,
                                           zinb_prob = zinb_prob, diff.pct = diff.pct, gene_params, seed = seed)
  
  srat_injected
})
data_S_merge_test[["umap"]] = data_S_merge[["umap"]]
```
```{r}
DimPlot(data_S_merge_test, group.by = "signal_cells",
        reduction = "umap", 
        na.value = "lightgrey", 
        cols = color_m,order = T, split.by = "batch", pt.size = 1.5) & NoAxes() & ggtitle(NULL)
```

# Figure 2B
```{r}
data_S_ls_test = SplitObject(data_S_merge_test,split.by = "batch")
data_S_ls_test = lapply(data_S_ls_test, function(data_S){
  data_S = data_S %>% NormalizeData() %>%
    FindVariableFeatures() %>%
    ScaleData(verbose = FALSE) %>% 
    RunPCA(npcs = 50, verbose = FALSE) 
  data_S
})
```
```{r}
# RunDUFS
common_genes = union(SelectCommonGenes(data_S_ls_test[[1]],data_S_ls_test[[2]],
                                   ngenes = 500),names(injected_genes))

lapply(1:length(data_S_ls_test),function(i){
  if(!file.exists(file.path(res.path,paste0("GeneTrajectory",i)))){
    data_S = data_S_ls_test[[i]]
    ComputeGeneEMD(data_S, common_genes, dir.path = file.path(res.path,paste0("GeneTrajectory",i)), script_dir = "../DiCoLo/vignettes")
  }
})

gene_emd_ls = lapply(1:length(data_S_ls_test),function(i){
  LoadGeneEMD(file.path(res.path,paste0("GeneTrajectory",i),""))
})

# Compute gene graph
gene_graph_ls = lapply(gene_emd_ls, ComputeGraphOperator)

# Visualize injected genes
set.seed(4)
other_genes = sample(setdiff(rownames(gene_emd_ls[[1]]),names(injected_genes)))[1:50]
order_levels = c("c_1", "c_4", "1_2", "1_3", "2_2", "2_3")
injected_genes_sorted <- factor(injected_genes, levels = order_levels)
injected_genes_sorted <- injected_genes_sorted[order(injected_genes_sorted)]
genes = c(names(injected_genes_sorted),other_genes)

pl <- lapply(1:2,function(i){
  gene_emd_dist = gene_graph_ls[[i]]
  mat <- gene_emd_dist[genes,genes]
  # hard-cap the matrix
  cap_value <- quantile(mat[upper.tri(mat, diag = FALSE)], 0.95)
  mat_capped <- mat
  mat_capped[mat_capped > cap_value] <- cap_value
  # min-max normalization
  mat_capped = (mat_capped-min(mat_capped))/(max(mat_capped) - min(mat_capped))
  
  p = VisualizeGeneHeatmap(mat = mat_capped, gene_partition = injected_genes_sorted, 
                       # gene_order = hclust(as.dist(gene_emd_dist[genes,genes]), method = "average")$'order',
                       gene_order = 1:length(genes),
                       # highlighted_genes = genes,
                       module_color = color_m[order_levels])
  as.ggplot(p)
})

wrap_plots(pl, nrow = 2) + plot_layout(guides = "collect")

```

# Figure 2C
```{r}
sample_diff = 1
diff.op = ComputeDifferentialOperator(gene_graph_ls[sample_diff][[1]],gene_graph_ls[-sample_diff][[1]])
E.list <- RunSVD(diff.op, eig_keep = nrow(diff.op))
n_eigvec = FindKneePoint(E.list$values[1:floor(sqrt(length(E.list$values)))],plot.fig = TRUE) - 1
```
```{r}
other_genes = setdiff(rownames(gene_emd_ls[[1]]),names(injected_genes))
order_levels = c("c_1", "c_4", "1_2", "1_3", "2_2", "2_3")
injected_genes_sorted <- factor(injected_genes, levels = order_levels)
injected_genes_sorted <- injected_genes_sorted[order(injected_genes_sorted)]
genes = c(names(injected_genes_sorted),other_genes)

pl = lapply(1:n_eigvec,function(i){
  p1 = Visualize_Eigenvec(E.list$vectors[genes,i],injected_genes_sorted)
})
wrap_plots(pl, nrow = 1)
```

```{r}
sample_diff = 2
diff.op = ComputeDifferentialOperator(gene_graph_ls[sample_diff][[1]],gene_graph_ls[-sample_diff][[1]])
E.list <- RunSVD(diff.op, eig_keep = nrow(diff.op))
n_eigvec = FindKneePoint(E.list$values[1:25],plot.fig = TRUE) - 1
```
```{r}
other_genes = setdiff(rownames(gene_emd_ls[[1]]),names(injected_genes))
order_levels = c("c_1", "c_4", "1_2", "1_3", "2_2", "2_3")
injected_genes_sorted <- factor(injected_genes, levels = order_levels)
injected_genes_sorted <- injected_genes_sorted[order(injected_genes_sorted)]
genes = c(names(injected_genes_sorted),other_genes)

pl = lapply(1:n_eigvec,function(i){
  p1 = Visualize_Eigenvec(E.list$vectors[genes,i],injected_genes_sorted)
})
wrap_plots(pl, nrow = 1)
```

# Figure 3
Code for generating benchmarking results `benchmarking_result.csv` for each dataset can be found in `run_benchmark.R`.
```{r}
dataset_ls = c("pbmc10k","pbmcsca","panc8")
dataset_full_name = setNames(c("Human pbmc10k","Human pbmcsca","Human pancreas"),dataset_ls)
method_full_name = setNames(c("DiCoLo","DGCA","LEMUR","miloDE"),c("DiCoLo","DGCA","lemur","milode"))
para_full_name = setNames(c("neighborhood size","number of genes","dropout rate"),c("ncell","ngene","dropout-rate"))
total_cell_ls = setNames(c(1108,596,1423),dataset_ls)
df = do.call(rbind,lapply(dataset_ls,function(dataset){
  tmp_path = dataset
  df = read.csv(file.path(dir.path, tmp_path, "benchmarking_result.csv"))
  df$'dataset' = dataset
  df
}))
df$method = factor(df$method,levels = names(method_full_name))
levels(df$method) = method_full_name[levels(df$method)]
df$'para_grid' = ifelse(df$para_test == "ncell",sprintf("%s(%d)",df$para_grid,floor(df$para_grid*total_cell_ls[df$dataset])),
                        df$'para_grid')
df$dataset = factor(df$dataset,levels = names(dataset_full_name))
levels(df$dataset) = dataset_full_name[levels(df$dataset)]
df$para_test = ifelse(df$para_test == "zinb_prob","dropout-rate",df$para_test)
df$para_test = para_full_name[df$para_test]

df_summary <- df %>%
  group_by(method, para_grid, para_test, dataset) %>%
  summarise(
    median_score = median(score, na.rm = TRUE),
    lower = pmax(0,median_score - sd(score, na.rm = TRUE) / sqrt(n())),  # Standard error (SE)
    upper = pmin(1,median_score + sd(score, na.rm = TRUE) / sqrt(n()))
  )

pl = lapply(unique(df_summary$para_test),function(para){
  if(para != "neighborhood size"){
    df_summary$para_grid = as.factor(as.numeric(df_summary$para_grid))
    angle = 0
  }else{angle = 90}
  ggplot(data = df_summary %>% filter(para_test == para), 
         aes(x = para_grid, y = median_score, 
             group = method, color = method,
             # linetype = method,shape = method
             )) +
    geom_line() +
    geom_point(size = 2) + 
    geom_errorbar(aes(ymin = lower, ymax = upper), 
                  width = 0.4,linewidth = 0.8) + 
    facet_grid( ~ dataset, scales = "free_x") +
    # scale_y_continuous(limits = c(0,1)) +
  labs(x = para, 
       y = "Normalized AUPRC", 
       color = "Method") + 
    theme(
      legend.title = element_text(size = 20),
      legend.text = element_text(size = 15),
      strip.text.x = element_text(size = 15),
      axis.title.x = element_text(size = 20),
      axis.title.y = element_text(size = 20),
      axis.text.x = element_text(size = 15,angle = angle),
      axis.text.y = element_text(size = 15),
      panel.grid = element_blank(),
      panel.background = element_blank(),
      axis.line = element_line(colour = "black"))
  
})
pl = lapply(1:length(pl),function(i){
  p = pl[[i]]
  if(i!=1){
    p = p + theme(strip.text.x = element_blank())
  }
  p = p + labs(y = NULL)
  p
})

wrap_plots(pl,ncol = 1) + plot_layout(guides = "collect")
```

# Figure 4A
The preprocessed Seurat object can be downloaded from [figshare for Mutant](https://figshare.com/ndownloader/files/55008218) and [figshare for Wildtype](https://figshare.com/ndownloader/files/55008224).
```{r}
sample_ls = paste0("smom2_dermal_",paste0("E13.5_",c("MUT","CTL")))
data_S_ls = lapply(sample_ls,function(sample_name){
  readRDS(file.path(dir.path.data,sprintf("data_S_%s.rds",sample_name)))
})
names(data_S_ls) = sample_ls

# Visualize Merged embedding -----
data_S_merge = merge(data_S_ls[[1]],data_S_ls[-1])
data_S_merge = data_S_merge %>% NormalizeData() %>% 
  FindVariableFeatures() %>% ScaleData() %>%
  RunPCA(npcs = 50, verbose = FALSE) %>% 
  RunUMAP(dims = 1:15,seed = 42)

DimPlot(data_S_merge, group.by = 'orig.ident', shuffle = TRUE) + NoAxes() + 
  ggtitle(NULL)
```

# Figure 4B
```{r}
color_ct = setNames(c("darkred","darkblue","#AC9362"),c("DC","LD","UD"))
pl = lapply(data_S_ls,function(data_S){
  DimPlot(data_S, group.by = "celltype", cols = color_ct) + NoAxes() + ggtitle(NULL)
})
(wrap_plots(pl,nrow = 2) & NoAxes() & ggtitle(NULL)) + plot_layout(guides = "collect")
```
```{r}
data.path = file.path(dir.path,"smom2")
merged_modules = do.call(c,lapply(c("MUT","CTL"),function(sample_name){
  readRDS(file.path(data.path,paste0(sample_name,"_dicolo_modules.rds")))
}))
merged_modules_ls = split(names(merged_modules), merged_modules)[1:4]
module_prefix = "Module_"
data_S_ls = lapply(data_S_ls, function(data_S){
  DefaultAssay(data_S) = "RNA"
  # data_S = AddModuleActivityScore(data_S, gene_partition = merged_modules)
  data_S <- AddModuleScore(
    object = data_S,
    features = merged_modules_ls,
    ctrl = 100,
    name = module_prefix
  )
  module_score = data_S@meta.data[,grepl(module_prefix,colnames(data_S@meta.data))]
  module_score = apply(module_score,2,function(x) scale(x)[,1])
  module_score = apply(module_score,2,function(x) (x - min(x)) / (max(x) - min(x)))
  data_S = AddMetaData(data_S,module_score)
})
pl = lapply(data_S_ls,function(data_S){
  FeaturePlot(data_S, features = paste0(module_prefix,seq_len(length(merged_modules_ls))), 
              # cols = colorRampPalette(rev(brewer.pal(n = 10, name = "RdBu")))(100),
              ncol = length(merged_modules_ls)) & 
  scale_colour_gradientn(colours = brewer.pal(n = 9, name = "Reds")) & NoAxes()
})
(wrap_plots(pl,nrow = 2) & NoAxes() & ggtitle(NULL)) + plot_layout(guides = "collect")
```


# Figure 4C
```{r}
library("ReactomePA")
library("org.Mm.eg.db")
library("clusterProfiler")
library("msigdbr")
m_df = msigdbr(species = "Mus musculus", category = "H")
m_df$gs_name = gsub("HALLMARK_","",m_df$gs_name)
msig_result <- do.call(rbind,lapply(names(merged_modules_ls), function(i){
  df = enricher(gene=merged_modules_ls[[i]],
           TERM2GENE = m_df[,c("gs_name","gene_symbol")],
           pAdjustMethod = "none",pvalueCutoff = 0.1)@result
  df$'modules' = i
  df = df %>% filter(p.adjust < 0.2)
  df = df[1:5,]
  df <- df[complete.cases(df), ]
  df
}))
msig_result$GeneRatio = unlist(lapply(msig_result$GeneRatio,function(x) eval(parse(text = x))))
msig_result$logP = -log(msig_result$pvalue,base = 10)
msig_result$Description <- factor(msig_result$Description,
                                 levels = rev(unique(msig_result$Description[order(msig_result$pvalue)])))

msig_result$modules = factor(msig_result$modules,levels = names(merged_modules_ls)[1:5])
levels(msig_result$modules) = gsub("MUT","SmoM2_M",levels(msig_result$modules))
levels(msig_result$modules) = gsub("CTL","CTL_M",levels(msig_result$modules))
# Dotplot
p = ggplot(msig_result, aes(x=modules, y=Description)) +
  geom_point(aes(size=GeneRatio, color=logP), alpha=0.7) +
  scale_color_gradientn(colors=colorRampPalette(c("blue","red"))(99), name="-log10_pvalue") +
  scale_size_continuous(name = "Gene Ratio", limits = c(0, 1)) + 
  theme_minimal() +
  labs(size="Gene Ratio", x=NULL,y=NULL) + 
  theme(#legend.position = "bottom",
        #legend.box = "vertical",
        axis.text.x = element_text(size = 10, angle = 45,hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 10)
  )
ggsave(file.path(figure.path, "fig4D_reactome.png"), p, 
       width = 8, height = 4)
```

# SupFigS2E
```{r}
# ---- Moran's I for a vector x under sparse weights W --------
# W: sparse adjacency (n x n). Row-standardized inside.
moran_I <- function(x, W) {
  x  <- as.numeric(x)
  xc <- x - mean(x)
  rs <- Matrix::rowSums(W)
  rs[rs == 0] <- 1                      # isolated cells -> no contribution
  Wn <- W / rs                          # row-standardize (W-style weights)
  S0 <- sum(Wn@x)
  num <- as.numeric(Matrix::crossprod(xc, Wn %*% xc))  # sum_ij w_ij xc_i xc_j
  den <- sum(xc^2)
  (length(x) / S0) * (num / den)
}
 
knn_W <- function(emb, k) {
  nn <- FNN::get.knn(emb, k = k)$nn.index          # n x k
  n  <- nrow(emb)
  W  <- Matrix::sparseMatrix(i = rep(seq_len(n), each = k),
                     j = as.vector(t(nn)),
                     x = 1, dims = c(n, n))
  as(W, "dgCMatrix")                               # row-standardised inside moran_I
}
 
module_moran_equalN <- function(data_S_ls,
                                module_prefix = "Module_",
                                dims = 1:10, k = 10,
                                frac = 0.9,      # subsample BOTH to frac*min(N) so both vary
                                R = 30, seed = 1) {
  # embedding + score matrix per condition (PCA built once if missing)
  prep <- lapply(data_S_ls, function(data_S) {
    DefaultAssay(data_S) <- "RNA"
    if (!"pca" %in% Reductions(data_S)) {
      data_S <- data_S |> NormalizeData(verbose = FALSE) |>
        FindVariableFeatures(verbose = FALSE) |> ScaleData(verbose = FALSE) |>
        RunPCA(npcs = max(dims), verbose = FALSE)
    }
    cols <- grep(module_prefix, colnames(data_S@meta.data), value = TRUE)
    list(emb   = Embeddings(data_S, "pca")[, dims, drop = FALSE],
         score = as.matrix(data_S@meta.data[, cols, drop = FALSE]),
         cols  = cols)
  })
 
  n_target <- floor(frac * min(sapply(prep, function(p) nrow(p$emb))))  # < min(N): both vary
  message("equal-N target = ", n_target, " cells per condition (", frac,
          " x min N); R = ", R)
 
  set.seed(seed)
  do.call(rbind, lapply(names(prep), function(cond) {
    p <- prep[[cond]]
    # R subsamples -> matrix of I values (R x modules)
    Imat <- t(replicate(R, {
      idx <- sample(nrow(p$emb), n_target)
      W   <- knn_W(p$emb[idx, , drop = FALSE], k)
      vapply(seq_along(p$cols), function(j) moran_I(p$score[idx, j], W), numeric(1))
    }))
    colnames(Imat) <- p$cols
    data.frame(
      module    = p$cols,
      condition = cond,
      I_mean    = colMeans(Imat),
      I_sd      = apply(Imat, 2, sd),
      I_lo      = apply(Imat, 2, quantile, 0.025),
      I_hi      = apply(Imat, 2, quantile, 0.975),
      n_target  = n_target, R = R, row.names = NULL)
  }))
}
 
moran_eq <- module_moran_equalN(data_S_ls, module_prefix = "Module_",
                                dims = 1:10, k = 10, R = 30)

# ---- equal-N paired plot (error bars = subsampling 95% interval) ----
moran_eq$module <- factor(moran_eq$module, levels = unique(moran_eq$module))
levels(moran_eq$module) = c("SmoM2_M1","SmoM2_M2",
                            "SmoM2_M3","CTL_M1")
moran_eq$condition = gsub("smom2_dermal_E13.5_","",moran_eq$condition)
 
p_eq <- ggplot(moran_eq, aes(x = module, y = I_mean, color = condition)) +
  geom_line(aes(group = module), color = "grey60", linewidth = 0.4,
            position = position_dodge(width = 0.5)) +
  geom_errorbar(aes(ymin = I_lo, ymax = I_hi), width = 0.15,
                position = position_dodge(width = 0.5)) +
  geom_point(position = position_dodge(width = 0.5), size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey70") +
  labs(x = NULL, y = "Moran's I", color = NULL
       ) +
  theme_classic(base_size = 13) +
  theme(legend.position = "top")
 
print(p_eq)

ggsave(file.path(figure.path, "figS2E.png"), p_eq, 
       width = 5, height = 4)
```

# Figure 4D
```{r}
sample_ls = paste0("smom2_dermal_",paste0("E13.5_",c("MUT","CTL")))
dir.path = "DiCoLo_data"
data.path = file.path(dir.path,"smom2")
res.path = file.path(data.path, "parameters")
gene_emd_ls = lapply(c("MUT","CTL"),function(i){
  LoadGeneEMD(file.path(res.path,paste0("GeneTrajectory_",i),""))
})

# Compute gene graph
gene_graph_ls = lapply(gene_emd_ls, ComputeGraphOperator)
merged_modules_sub = stack(merged_modules_ls) |> setNames(c("gene", "group"))
merged_modules_sub = setNames(merged_modules_sub$group,merged_modules_sub$gene)
genes = names(merged_modules_sub)
color_m = setNames(brewer.pal(4,"Set1")[1:length(unique(merged_modules_sub))],unique(merged_modules_sub))

pl <- lapply(1:2,function(i){
  gene_emd_dist = gene_graph_ls[[i]]
  mat <- gene_emd_dist[genes,genes]
  # hard-cap the matrix
  cap_value <- quantile(mat[upper.tri(mat, diag = FALSE)], 0.95)
  mat_capped <- mat
  mat_capped[mat_capped > cap_value] <- cap_value
  # min-max normalization
  mat_capped = (mat_capped-min(mat_capped))/(max(mat_capped) - min(mat_capped))
  
  p = VisualizeGeneHeatmap(mat = mat_capped, gene_partition = merged_modules_sub, 
                       # gene_order = hclust(as.dist(gene_emd_dist[genes,genes]), method = "average")$'order',
                       gene_order = 1:length(genes),
                       # highlighted_genes = genes,
                       module_color = color_m)
  as.ggplot(p)
})

p = wrap_plots(pl, nrow = 2) + plot_layout(guides = "collect")
ggsave(file.path(figure.path, "fig4D_heatmap.png"), p, 
       width = 4, height = 6)
```

